---
title: "Geometrical Anatomy of Theoretical Physics Lecture 2"
description: |
  Set Theory
author:
  - name: Phu Sakulwongtana
    url: https://phutoast.github.io/
date: 01-21-2019
output:
  radix::radix_article:
    self_contained: false
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The $\epsilon$-Relation
Set Theory is built on the postualte that there is a fundamental relation caleed 

$$
\in
$$

There will be no definition of what $\in$ is or what a set is. Instead, 9 axioms that use of $\in$ and set, will show hows they works together. 

Using $\in$-relation, we can immediately define 3 more operations 

$$
\begin{aligned}
\not\in (x, y) \equiv x \not\in y &:\iff \neg(x \in y) \\
x \subseteq y &:\iff \forall a : (a \in x \implies x \in y) \\ 
x = y &:\iff x \subseteq y \wedge y \subseteq x
\end{aligned}
$$

---

## ZFC Set Theory

### Axiom of $\in$-relation 

This is proposition if and only if $x$ and $y$ are set 
$$
x \in y
$$

The counter example of this axiom would be (Russell's Paradox) -- Assume there is a some set $u$ that contains all set that doesn't contain themselves. To be precise 

$$
\exists u : \forall z (z \in u \iff z \not\in z)
$$

Question -- Is $u$ a set ? 

  * If $u$ was a set, then one must be able to decide whether, $u \in u$ is true or false. 
  * But assume, that $u \in u$ then by definition of $u$, $$u \not\in u$$

Assume that 

$$
u \in u\text{ is false } \iff u \not\in u \implies u \in u
$$

The Conclusion -- $u$ is not a set, a *Contradiction*. Having this one axiom is not enought.

### Axiom of Existence of An Empty Set

There exists a set that contains no elements 

$$
\exists x : \forall y : y \not\in x
$$

For the more formal version.

$$
\forall x : \forall y : (x\in y) \oplus \neg(x \in y)
$$

__Theorem__ -- There is only one empty set, we call it $\emptyset$

__Proof (Normal Style)__ -- Assume $x'$ and $x$ are both empty set, then 

$$
\forall y : (y \in x) \implies (y \in x')
$$

But this is equivalent to $x \subseteq x'$. Conversely 

$$
\forall y : (y \in x') \implies (y \in x)
$$

So, $x'\subseteq x$, which means $x = x'$

__Proof (Formal Version)__ -- There will only be one formal proof in this course.

$$
\begin{aligned}
\text{(Axiom) } a_1 &\iff \forall y : y \not\in x \\ 
\text{(Axiom) } a_2 &\iff \forall y : y \not\in x' \\ 
\\
\text{(Tautology) } q_1 &\iff \forall y \not\in x \implies \forall y : (y \in x \implies y \in x') \\ 
\text{(Axiom) } q_2 &\iff \forall y : y \not\in x \\ 
\text{(Modus Ponens) } q_3 &\iff \forall y : (y \in x \implies y \in x') \iff x \subseteq x' \\
\text{(Tautology) } q_4 &\iff \forall y \not\in x' \implies \forall y : (y \in x' \implies y \in x) \\ 
\text{(Axiom) } q_5 &\iff \forall y : y \not\in x' \\ 
\text{(Modus Ponens) } q_6 &\iff \forall y : (y \in x' \implies y \in x) \iff x' \subseteq x \\
\text{(Modus Ponens) } q_7 &\iff x = x'
\end{aligned}
$$

### Axiom of Pair Sets

Let $x$ and $y$ be sets then there exists a set as its element precisely the set $x$ and $y$.

$$
\forall x \forall y : \exists m : \forall u (u \in m \implies u =x \vee u = y)
$$

Notation -- We denote this $m$ by $\{x, y\}$, also we can see that $\{x, y\} = \{y, x\}$, can be proven by 

$$
\begin{aligned}
a \in \{x, y\} &\implies a \in \{y, x\} \\
a \in \{y, x\} &\implies a \in \{x, y\}
\end{aligned}
$$

Which means 
$$
\{x, y\} \subseteq \{y, x\} \text{ and } \{y, x\} \subseteq \{x, y\} \iff \{x, y\} = \{y, x\}
$$

__Remark, Definition__ -- $\{x\} = \{x, x\}$

### Axiom of Union Set

Let $x$ be a set, then there exists a set $u$ whose element are preciselt the elements of the element of x. Notation -- 

$$
u = \bigcup x
$$

For example, Let $a$ and $b$ be sets 

$$
\{\{a\}, \{b\}\} \text{ then } \bigcup x = \{a, b\}
$$

__Definition__ -- Let $a_1, a_2, ..., a_N$ be a sets defined recursively, then 

$$
\{a_1, ..., a_N\} = \bigcup \left\{ \{a_1, ..., a_{N-1}\}, \{a_N\} \right\} \text{ for all } n \ge 3
$$

### Axiom of Replacement

Let $R$ be a functional relation. Let $m$ be a set. Then the image of $m$ under the functional relation is again a set. 

Here -- A relation $R$ is called a function if 

$$
\forall x \exists ! y : R(x, y) \text{ (There exists precisely one) }
$$

The image of a set $m$ under a functional relation $R$ consists of all those $y$ for which there is an $x \in m$, such that $R(x, y)$

*Axiom of replacement* implies, but isn't implied by the *Principle of Restricted Comprehension*. Let $p$ be a predicate of one variable and let $m$ be a set. Then those element $y \in m$ for which $p(y)$ holds, constitute a set. Notation -- This set is usually denote as 

$$
\{y \in m | p (y)\}
$$

*Principle of Restricted Comprehension* isn't to be confused with the inconsistent principle of universal comprehension 

$$
\{y | p(y)\} \text{ is not a set}
$$
You will need to tell which set it comes from (have to be bounded by set $m$). 

__Proof__ (That *Principle of Restricted Comprehension*  is implied by *Axiom of replacement*) - Case 1.

$$
\neg \exists y \in m : P(y)
$$

In this case -- 

$$
\{y \in m | P(y)\} = \emptyset
$$

Case 2 -- 
$$
\exists \hat{y} \in m : P(\hat{y})
$$

Then we define 
$$
R(x, y) := (P(x) \wedge x = y) \vee (\neg P(x) \wedge \hat{y} = y)
$$
we claim that $R$ is a function.

__Definition__ -- Let 

$$
u \subseteq m \text{ then } m \backslash u := \{x \in m | x \not\in u \}
$$

It is a set due to *Principle of Restricted Comprehension*, ultimately due to axiom of replacement. 

### Axiom of Existence of Power Set

Historically, in naive set theory, *Principle of Restricted Comprehension* was throught to be needed in and to define, for any set $m$

$$
\mathcal{P}(m) = \{y | y \subseteq m\}
$$

but it is inconsistence with *Principle of Restricted Comprehension*, so let $m$ be a set. Then there exists a set denote $\mathcal{P}(m)$ whose element are precisely the subset of $m$, for example

$$
\mathcal{P}(m) = \{\emptyset, \{a\}, \{b\}, \{a, b\}\}, \text{ where } m = \{a, b\}
$$

### Axiom of Infinity 

There exists a set that contains the empty set and with every of its element $y$ is also contains \{y\} as an element. 

<aside>
This is an old definition. 
</aside>

Remark -- One such set is informally speaking the set with element 
$$
\{\emptyset, \{\emptyset\}, \{\{\emptyset\}\}, ...\}
$$

Corollary -- $\mathbb{N}$ is a set, and as a set 

$$
\mathbb{R} = \mathcal{P}(\mathbb{N})
$$

### Axiom of Choice

Let $x$ be a set whose element are 

* Non-Empty 
* Mutuallly Disjoint 

Then there exists a set $y$ with contains exactly one element of each element of $x$. *Remark* -- Set $y$ is called "darkset"

__Intuition__ 

$$
\begin{aligned}
x &= \left\{ \{\text{Left Shoe 1}, \text{Right Shoe 1}\}, \{\text{Left Shoe 2}, \text{Right Shoe 2}\}, ... \right\} \\
y &= \left\{ \text{Left Shoe 1}, \text{Left Shoe 2}, ... \right\}
\end{aligned}
$$

This doesn't need axiom of choice. 
$$
\begin{aligned}
x &= \left\{ \{\text{Sock 1 (Left)}, \text{Sock 1 (Right)}\}, \{\text{Sock 2 (Left)}, \text{Sock 2 (Right)}\}, ... \right\} \\
y &= \left\{ \text{Sock 1}, \text{Sock 2}, ... \right\}
\end{aligned}
$$

*Remark* -- Axiom of choice is independency of the other 8 axioms. We could have a set theory with or without it. 

*Remark* -- Proof that every vector space has a basis need axiom of choice. Proof that there exists a complete system of representation of an equivalent relation, needs Axiom of choice

### Axiom of Foundation 

Every non-empty set $x$ contains an element $y$ that has none of its element in common with $x$. Immdiate Implication is that, there is no set that contains itself as an element.

```{r, echo=FALSE}
htmltools::includeHTML("katex.html")
```