---
title: "Geometrical Anatomy of Theoretical Physics Lecture 3"
description: |
   Classification of Sets, Construction of of ℕ, ℤ and ℚ 
author:
  - name: Phu Sakulwongtana
    url: https://phutoast.github.io/
date: 01-22-2019
output:
  radix::radix_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

---

## Classification of Sets

### Maps
A recurrence theme in mathematics is the study/classification of spaces by means of structure preserving maps between these spaces. A space is usually meant to be same equipped with some structure. 

A map 
$$
\phi : A \rightarrow B
$$

is a relation such that for every $a \in A$ there exists exactly one $b \in B$ such that $\phi(a, b)$

__Notation__ -- 

$$
\phi : A \rightarrow B \quad \phi (a) :=  b
$$

We say that 

* $A$ is a domain of $\phi$
* $B$ is a target of $\phi$
* Image is defined by 

$$
\phi(A) \equiv\text{im}_{\phi} (A) := \left\{ \phi(a) \in B | a \in A \right\}
$$

__Definition__ -- A map 

$$
\phi: A \rightarrow B 
$$
is called 

* Subjective if $\phi(A) = B$
* Injective if $\phi(a_1) = \phi(a_2) \implies a_1 = a_2$
* Bijective if the map is both injective and surjective. 

__Definition__ -- Two sets $A$ and $B$ are called (Set-Theoretically) isomorphic 

$$
A \cong_{\text{set}} B
$$

If there exists a bijection 

$$
\phi : A \rightarrow B
$$

__Remark__ -- If there is any bijection $A \rightarrow B$ then generally there are many. (The intuition for this is that we pair up elements of $A$ and $B$)

### Classification of Set

A set $A$ is 

* Infinite if there exits a propert subset $B$. $B \subsetneqq A$. That is isomorphic to A, $B \cong_{\text{set}} A$

  * $A$ is called countably infinite if $A \cong_{\text{set}} \mathbb{N}$ 
  * $A$ is called non-countably otherwise 

* Finite, otherwise, or we can say that 

$$
A \cong_{\text{set}} \{1, 2, ..., N\} \text{ For Some } N \in \mathbb{N}
$$
Where we write 

$$
|A| = N
$$

__Definition__ - Let a map be $\phi : A \rightarrow B$ be a bijection. Then the inverse of the map is a map $\phi^{-1} : B \rightarrow A$ defined uniquely by 

$$
\begin{aligned}
\phi \circ \phi^{-1} &= \text{id}_A \\
\phi^{-1} \circ \phi &= \text{id}_B
\end{aligned}
$$
Where $\text{id}_M : M \rightarrow M \text{ where } m \mapsto m$

__Definition__ -- Let the map $\phi: A \rightarrow B$ by an map. Then we define the set 

$$
\text{pre-im}_{\phi} (V) = \{a \in A | \phi(a) \in V\} \text{ where } V \subseteq B
$$

### Equivalent Relation 

__Definition__ -- Let $M$ be a set and $\sim$ a relation such that 

* It has the property of reflexivity 

$$
\forall m \in M : m \sim m
$$

* It has the property of symmetry 

$$
\forall m, n \in M : m \sim n \iff n \sim m
$$

* It has the property of transitivity 

$$
\forall m, n, p \in M : m \sim n \wedge n \sim p \implies m \sim p
$$

Then $\sim$ is called equilvalent relation on $M$

__Definition__ -- Let $\sim$ be an equivalent relation on $M$. Then for any $m \in M$. We define a set 

$$
[m] = \{n \in M | m \sim n\} \subseteq M 
$$

There are 2 key properties 

$$
a \in [m] \implies [a] = [m]
$$

and it is either 

$$
[m] = [n] \text{ or } [m] \cap [n] = \emptyset
$$


__Definition__ -- Let $\sim$ be an equivalent relation on $M$. Define the quotient set 

$$
M / \sim = \{ [m] \in \mathcal{P}(M) | m \in M \}
$$

__Remark__ -- Due to the axiom of choice, there exists a complete system of representation for $\sim$ in the set $R$ such that 

$$
R \cong_{\text{set}} M / \sim 
$$

__Remark__ -- Care must be taken to define map whose domain is a quotient set and if one uses representative to define the map 

__Example__ -- Suppose 

$$
M = \mathbb{Z} , \quad m \sim n \iff m-n \in 2\mathbb{Z}
$$

Then the quotient set would be 

$$
\mathbb{Z} / \sim = \{[0], [1]\}
$$

Idea -- On $\mathbb{Z}$ we have addition 

$$
+ : \mathbb{Z} \times \mathbb{Z} \rightarrow \mathbb{Z}
$$

We wish to inherit and addition as 

$$
\oplus : \mathbb{Z/\sim  \times \ Z /\sim \ \rightarrow \ \mathbb{Z} /\sim }
$$

Such that 

$$
[a] \oplus [b] = [a+b]
$$

Care must be taken precisely because this could be inconsistent, we have to show that, let $a'$ and $b'$, such that 

$$
[a'] = [a], \ [b'] = [b]
$$

Then we can see that 

$$
\begin{aligned}
[a'] \oplus [b'] &= [a' + b'] = [a-2m + b - 2n] \\
&= [a + b - 2(m+n)] = [a+b] = [a] \oplus [b]
\end{aligned}
$$
And indeed it is well-defined.

---

## Construction of $\mathbb{N, Z}$ and $\mathbb{ Q }$

### Natural Number $\mathbb{N}$

We would like to establish an addition on $\mathbb{N}$.

<aside>
This is based on an old definition of $\mathbb{N}$
</aside>

__Define__ -- Successor Map, such that 

$$
s : \mathbb{N} \rightarrow \mathbb{N} \quad n \mapsto \{n\}
$$

For example , 

$$
s(2) = s(\{\{\emptyset\}\}) = \{\{\{\emptyset\}\}\} = 3
$$

__Define__ -- Predecessor Map 

$$
p : \mathbb{N}^* \rightarrow \mathbb{N} \quad n \mapsto m \text{ such that } m \in n
$$

where $\mathbb{N}^*$ is a natural network that doesn't contain $0$. For example. 

$$
p(2) = p(\{\{ \emptyset \} \}) = \{\emptyset\} = 1
$$

__Define__ -- n-th power of $s$ means that 

$$
s^n = s \circ s^{p(n)} \text{ if } n \in \mathbb{N}^* \quad s^0 = \text{id}
$$

__Definition__ - Addition is a map 

$$
\mathbb{N} \times \mathbb{N} \rightarrow \mathbb{N} \quad (m ,n) \mapsto m + n = s^n(m)
$$

For example, 

$$
1 + 2 = s^2(1) = s(s(1)) = s(2) = 3 \quad \quad \quad 2 + 1 = s^1(2) = s(2) = 3
$$

We can show that it is commutative and we can also show that it is associative. 

### Integer Numbers $\mathbb{Q}$

It is defined by 
$$
\mathbb{Z = (\mathbb{N} \times \mathbb{N})} / \sim
$$

Where we define the equivalent relation as 

$$
(m, n) \sim (p, q) :\iff m+p = p + n
$$

And we can proof that it is an equivalent relation. The idea is that $(m, n)$ correspond for $m-n$.

__Comparison Between Integer and Natural Number__

$$
\mathbb{N \not\subseteq Z} \text{ because they have  difference structure}
$$
but 

$$
\mathbb{N \hookrightarrow Z} \quad n \mapsto [(n, 0)]
$$

__Definition__ -- Negative Number can be described as 

$$
n \in \mathbb{N} \quad -n = [(0, n)] \in \mathbb{Z}
$$

__Definition__ -- The addition for the integer numbers is defined by 

$$
+_{\mathbb{Z}} : \mathbb{Z \times Z \rightarrow Z}
$$

is defined as 

$$
[(m, n)] +_{\mathbb{Z}} [(p, q)] = [(m + p, n + q)]
$$

We can show this is well-defined. Let's try adding 2 given numbers 

$$
2 +{\mathbb{Z}} (-3) = [(2, 0)] +_{\mathbb{Z}} [(0, 3)] = [(2 + 0, 0 + 3)] = [(2, 3)] = -1
$$

### Rational Numbers

Rational Numbers is defined by 

$$
\mathbb{Q := (\mathbb{Z} \times \mathbb{Z}^*) / \sim}
$$

Now the equivalent relation is defined by 

$$
(x, y) \sim (u, v) :\iff xv = uy
$$

The idea behind this is that 

$$
\frac{a}{b} = [(x, y)]
$$

We can embedded $\mathbb{Z}$ by 

$$
\mathbb{Z \hookrightarrow Q}  \quad  x \mapsto[(x, 1)]
$$

__Define__ -- Addition on Rational Number as 

$$
+_{\mathbb{Q}} : \mathbb{Q \times Q \rightarrow Q}  
$$

where it is defined as 

$$
[(x, y)] +_{\mathbb{Q}} [(u, v)] =  [(x \cdot v + y \cdot u, y \cdot v)]
$$

Typical ill-defined addition would be 

$$
[(x, y)] +_{\mathbb{Q}} [(u, v)] =  [(x + u, y + v)]
$$

__Define__ -- The multiplication as 

$$
\cdot_{\mathbb{Q}} : \mathbb{Q \times Q \rightarrow Q}
$$

where 

$$
[(x, y)] \cdot_{\mathbb{Q}} [(u, v)] = [(x \cdot u, u \cdot v)]
$$

```{r, echo=FALSE}
htmltools::includeHTML("katex.html")
```
